---
title: "kyadav1_BA_Assignment1"
author: "Khushboo Yadav"
date: "10/18/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading Data and Importing libraries for Online Retail dataset

```{r}
Online_Retail <- read.csv("~/Desktop/MSBA/Business Analytics/Data/Online_Retail.csv")
View(Online_Retail)

library(dplyr)
library(caret)
library(tidyselect)
library(lubridate)
library(tidyverse)
library(Lahman)

str(Online_Retail) #Analyzing the structure of the dataset
nrow(Online_Retail) #Determining the row count of the dataset
```
## 1.Show the breakdown of the number of transactions by countries i.e. how many transactions are in the dataset for each country (consider all records including cancelled transactions). Show this in total number and also in percentage. Show only countries accounting for more than 1% of the total transactions. (5 marks)
```{r}
#Breakdown of the no. of transactions  in numbers and in percentage by Countries

TransactByCountry<-Online_Retail %>% 
  group_by( Country ) %>% 
  summarise( Percentage = 100 * (n() / nrow( Online_Retail ) ),Number_of_Transactions=n()  ) %>%
  
#Printing the complete results
print(TransactByCountry)

# Output : Countries accounting for for more than 1 percentage

filter(TransactByCountry,TransactByCountry$Percentage>=1)

```
## 2.Create a new variable ‘TransactionValue’ that is the product of the exising ‘Quantity’ and ‘UnitPrice’ variables. Add this variable to the dataframe. (5 marks)

```{r}
#TransactionValue = Quantity*UnitPrice
TransactionValue<-Online_Retail$Quantity*Online_Retail$UnitPrice

#Binding TransactionValue in Online Retail Dataframe
Online_Retail<-cbind(Online_Retail,TransactionValue)

#Using str function to validate the new addition in the Online Retail dataframe
str(Online_Retail)
```

## 3. Using the newly created variable, TransactionValue, show the breakdown of transaction values by countries i.e. how much money in total has been spent each country. Show this in total sum of transaction values. Show only countries with total transaction exceeding 130,000 British Pound. (10 marks)

```{r}
# Aggregating the results of Transaction Values based on different countries in the Online Retail Dataframe
Sum_TransactByCountry <- Online_Retail%>%
  group_by(Country)%>%
  summarise(Sum_Transaction_Value=sum(TransactionValue))


# Output:Filtering the results based on the 130000 British Pound
filter(Sum_TransactByCountry,Sum_Transaction_Value>130000)

```
## Q4.This is an optional question which carries additional marks (golden questions). In this question, we are dealing with the InvoiceDate variable.

```{r}
#explicitly instructing R to interpret InvoiceDate as a Date variable
Temp=strptime(Online_Retail$InvoiceDate,format='%m/%d/%Y %H:%M',tz='GMT')
head(Temp)
Online_Retail$New_Invoice_Date <- as.Date(Temp)
#convert dates to days of the week
Online_Retail$Invoice_Day_Week= weekdays(Online_Retail$New_Invoice_Date)
#Taking the hour from Temp 
Online_Retail$New_Invoice_Hour = as.numeric(format(Temp, "%H"))
#Taking months from Temp
Online_Retail$New_Invoice_Month = as.numeric(format(Temp, "%m"))
```
## 4.a) Show the percentage of transactions (by numbers) by days of the week (extra 2 marks)
```{r}
# Aggregating and determining percentage  of Transaction Values by number based on different countries
 Prcntg_N_DOW <- Online_Retail%>%
  group_by(Online_Retail$Invoice_Day_Week)%>%
  summarise(Transaction_DOW = (n()),PercentageTransaction = (n()/nrow(Online_Retail))*100)

# Output:To print the results
print(Prcntg_N_DOW)
```
                      
## 4.b) Show the percentage of transactions (by transaction volume) by days of the week (extra 1 marks)

```{r}
# Aggregating and determining percentage  of Transaction Values by volume based on different countries
 Prcntg_V_DOW<- Online_Retail%>%
  group_by(Online_Retail$Invoice_Day_Week)%>%
  summarise(Sum_TransactDOW = (sum(TransactionValue)))

#Calculating the percentage 
 Prcntg_V_DOW$PercentageTransaction<- (Prcntg_V_DOW$Sum_TransactDOW/sum(Prcntg_V_DOW$Sum_TransactDOW))*100 
 
# Output:Print the result
 print(Prcntg_V_DOW)

```

## 4.c) Show the percentage of transactions (by transaction volume) by month of the year (extra 1 marks)

```{r}
# Aggregating and determining percentage  of Transaction Values by volume based on different countries
Prcntg_MOY <- Online_Retail%>%
  group_by(Online_Retail$New_Invoice_Month)%>%
  summarise(Sum_TransactionValue = (sum(TransactionValue)))

#Calculating the percentage 
  Prcntg_MOY$Percentage_TransactionValue<- (Prcntg_MOY$Sum_TransactionValue/sum(Prcntg_MOY$Sum_TransactionValue))*100 
  
 # Output:Print the result
  print(Prcntg_MOY)
  
```
## 4.d) What was the date with the highest number of transactions from Australia? (3 marks)

```{r}
#Filtering the data based on Country = "Australia"
  Australia_Tran<- filter(Online_Retail,Online_Retail$Country=="Australia")

#Aggregating the transactions from Australia based on Invoice Dates
  Australia_Tran_bydate <- Australia_Tran%>%
                           group_by(Australia_Tran$New_Invoice_Date)%>%
                           summarise(TranCount_Aust = n())
  
 
  #Arranging the results in order to see the date with highest number of transactions 
  arrange(Australia_Tran_bydate,desc(Australia_Tran_bydate$TranCount_Aust))
  
#Output:Print the date with highest number of transactions  
 top_n(Australia_Tran_bydate,1)
```  
## 4.e) The company needs to shut down the website for two consecutive hours for maintenance. What would be the hour of the day to start this so that the distribution is at minimum for the customers? The responsible IT team is available from 7:00 to 20:00 every day. (3 marks)

```{r}
#Aggregating the no.of transactionValues by hours
  Tran_byhour <- Online_Retail%>%
    group_by(Online_Retail$New_Invoice_Hour)%>%
              summarise(TranCount = n())
             
 #Creating a variable with numeric data 
  smtrn <-numeric()
  
  #Using for loop for adding consecutive hours together:-
  
  for (i in 1:dim(Tran_byhour)[1])
   { 
  if (between(Tran_byhour$`Online_Retail$New_Invoice_Hour`[i],6,20)) 
    { 
      #Summation of the row [i] with row[i+1]
      smtrn<- c(smtrn,sum((Tran_byhour$TranCount[i]),(Tran_byhour$TranCount[i+1]))) 
     
    }
  }
 #Adding column
  Tran_byhour$NextHour<-Tran_byhour[,2]
  Tran_byhour$NextHour<-c(7:20,NA)
  
  #Combining the data into a single data frame
  Tran_byhour$ConsecutiveHours_CountTransaction<-smtrn
  
 #Printing the output
 print(Tran_byhour)
 
 
  
##Filtering 6 hours , due to unavailability of responsible IT team
Tran_byhour<-filter(Tran_byhour,Tran_byhour$`Online_Retail$New_Invoice_Hour`>6)
 

  #Output:determining the smallest(sum of transactions for2 consecutive hours) number 
top_n(Tran_byhour,-1)


```
 Answer:
Therefore the we can see the least no. of  consecutive transactions are happening in the 19-20 hrs.Hence 19hrs and 20 hrs are    best for doing the maintainence.


## 5. Plot the histogram of transaction values from Germany. Use the hist() function to plot. (5 marks)

```{r}
#Making a subset for Germany Transactions
Germany_TransactionValue<-subset(Online_Retail,Country=='Germany')

#To understand the data frame for a better Histogram
summary(Germany_TransactionValue)

ggplot(Germany_TransactionValue, aes(x=TransactionValue)) + 
  geom_histogram(breaks=seq(-600,900,by=20),col="Blue",fill="Green")+
  labs(title="Transaction Values in Germany",x="Transaction Values",y="Count")

```


## 6. Which customer had the highest number of transactions? Which customer is most valuable (i.e. highest total sum of transactions)? (10 marks)
## 6.1 Which customer had the highest number of transactions
```{r}
# Customers id  with the  number of transactions each
Customer_TranNo <- Online_Retail%>%
  group_by(CustomerID)%>%
  summarise( TransactionNo= n())

#arranging data in descending format to find the customer with maximum no. of transactions
 arrange(Customer_TranNo,desc(TransactionNo))

#removing na values from the dataset( highest no. of transactions are coming for the Customer 'IDs as "NA")
Customer_TranNo<-na.omit(Customer_TranNo)

#Output:populating the customer with highest no. of transaction after removing NA customer ID's
top_n(Customer_TranNo,1)
```

## 6.2 Which customer is most valuable (i.e. highest total sum of transactions)?
```{r}
#Valuable Customer based on the highest total sum of transaction
Customer_TranVol <- Online_Retail%>%
  group_by(CustomerID)%>%
  summarise(sum_Transaction = sum(TransactionValue))

#arranging data in descending format to find the customer ID with  highest total sum of transactions 
arrange(Customer_TranVol,desc(sum_Transaction))

#removing na values from the dataset(Since custmer id=NA  has the highest total sum of transactions)
Customer_TranVol<-na.omit(Customer_TranVol)

#Output:populating most valuable customer
top_n(Customer_TranVol,1)

```

## 7. Calculate the percentage of missing values for each variable in the dataset (5 marks). Hint colMeans():

```{r}
#Determining the Percentage of missing values for all the variables 
colMeans(is.na(Online_Retail))
```
## 8. What are the number of transactions with missing CustomerID records by countries? (10 marks)
```{r}
# Aggregating the date based on the countries and validating the customerID in terms of TRUE/FALSE by countries
 misCust <- Online_Retail%>%
  group_by(Country)%>%
  summarize(miss_val= is.na(CustomerID))

#Filtering the records based on missing values
mis_cust_cntry <- filter(misCust,miss_val=="TRUE")

#Output:Result for no. of transactions with missing CustomerID records by countries
count(mis_cust_cntry)

```  


## 9. On average, how often the costumers comeback to the website for their next shopping? (i.e. what is the average number of days between consecutive shopping) (Optional/Golden question: 18 additional marks!) Hint: 1. A close approximation is also acceptable and you may find diff() function useful.

```{r}

#Creating subset with only required columns

Online_Retail_2<-subset(Online_Retail[,c(1,4,5,7,10)])
Online_Retail_2 <- na.omit(Online_Retail_2) 

 #Aggregating the data based on Customer ID and determining the distinct InvoiceNo with the Invoice dates for all non-canceled transactions 

InvoiceDt_byCustomerID <- Online_Retail_2 %>% 
                         filter(Quantity>1) %>%
                         group_by(CustomerID) %>%
                         distinct(InvoiceNo,.keep_all = TRUE)%>%
                        arrange(CustomerID)

#Determining the difference of Invoice Dates
 Difference_InvoiceDt<-difftime(InvoiceDt_byCustomerID$New_Invoice_Date,lag(InvoiceDt_byCustomerID$New_Invoice_Date),units="days")
 

# Changing the type 
Difference_InvoiceDt<-as.data.frame((Difference_InvoiceDt))

Difference_InvoiceDt$Dif_InvoiceDT<-Difference_InvoiceDt

#Merging the Invoice date difference  with  the data frame using cbind
 InvoiceDt_byCustomerID<-data.frame(cbind(InvoiceDt_byCustomerID,Difference_InvoiceDt$Dif_InvoiceDT))


 
 ## To determine the consumer's come back, we don't need the customer's with no recurring transactions.Therefore we should remove them in order to get the right result.
 
# Updating the first record of difference in dates for  each customer as NA , as it was difference of last transaction  of one customer id with first transaction date of another.Also, this will help to remove the records where customer didn't come for the second purchase.
 
 InvoiceDt_byCustomerID <- InvoiceDt_byCustomerID %>%
                          group_by(CustomerID) %>%
                          mutate(X.Difference_InvoiceDt. = replace ( X.Difference_InvoiceDt. , row_number(CustomerID) == 1, NA))

#Removing the NA values 
InvoiceDt_byCustomerID<- na.omit(InvoiceDt_byCustomerID)

 #Output
 mean(InvoiceDt_byCustomerID$X.Difference_InvoiceDt.)
```

    
## 10. In the retail sector, it is very important to understand the return rate of the goods purchased by customers. In this example, we can define this quantity, simply, as the ratio of the number of transactions cancelled (regardless of the transaction value) over the total number of transactions. With this definition, what is the return rate for the French customers? (10 marks). Consider the cancelled transactions as those where the ‘Quantity’ variable has a negative value.

```{r}

#Filtering the records based on French customer with canceled transactions
France_cancelled<-filter(Online_Retail,Country=="France"&Quantity<0)


# Filtering the French Customer transactions
France_Total_Transactions<-filter(Online_Retail,Country=="France")


#Determining the rate of the canceled transactions by French Customer
rate<- (nrow(France_cancelled)/nrow(France_Total_Transactions))*100

#Output
print(rate)
```

## 11. What is the product that has generated the highest revenue for the retailer?(i.e. item with the highest total sum of ‘TransactionValue’). (10 marks)
      
```{r}     
#Aggregating the Transaction Value based on Product   
high_Revenue_Product <- Online_Retail%>%
  group_by(Online_Retail$Description)%>%
  summarise(highRev = sum(TransactionValue))

# Output:To determine the product with highest revenue
top_n(high_Revenue_Product,1)
```
## 12. How many unique customers are represented in the dataset? You can use unique() and length() functions. (5 marks)

```{r}
#Storing the unique customer ids in a new data frame
unique_customers<-data.frame(unique(Online_Retail$CustomerID,incomparables =FALSE,nmax=NA))

# Output:To check the count of unique customers
count(unique_customers)

```        
